---
title: "NYPD Shooting Incident Data Analysis"
author: "Cristian"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction
This report analyzes the NYPD Shooting Incident Data Historic dataset. It includes data on shooting incidents across various boroughs over time. The analysis involves data cleaning, visualizations, and a logistic regression model to predict shooting incidents.

## Load Libraries
```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(broom)
```

## Load and Clean Data
```{r}
nypd_data <- read_csv("NYPD_Shooting_Incident_Data__Historic_ (1).csv")

nypd_data_clean <- nypd_data %>%
  drop_na() %>%
  distinct() %>%
  mutate(
    OCCUR_DATE = mdy(OCCUR_DATE),
    BORO = factor(BORO),
    STATISTICAL_MURDER_FLAG = as.factor(STATISTICAL_MURDER_FLAG)
  )
```

## Visualization 1: Shooting Incidents by Borough
```{r}
ggplot(nypd_data_clean, aes(x = BORO)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Shooting Incidents by Borough", x = "Borough", y = "Number of Incidents") +
  theme_minimal()
```

## Visualization 2: Shooting Incidents Over Time
```{r}
ggplot(nypd_data_clean, aes(x = OCCUR_DATE)) +
  geom_histogram(binwidth = 30, fill = "lightgreen") +
  labs(title = "Shooting Incidents Over Time", x = "Date", y = "Number of Incidents") +
  theme_minimal() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
```

## Predictive Model: Logistic Regression
```{r}
model <- glm(STATISTICAL_MURDER_FLAG ~ BORO + OCCUR_DATE, data = nypd_data_clean, family = binomial())
summary(model)
```

## Bias Considerations
- **Temporal Bias**: Shooting incidents may be more prevalent in certain years or months.
- **Demographic Bias**: Certain racial groups may be overrepresented due to systemic factors.
- **Reporting Bias**: Data may not capture unreported incidents.

## Conclusion
This report explored trends in NYPD shooting incidents using visualizations and logistic regression modeling. Further research could incorporate additional predictive factors such as location and time of day.
